Fixing Vulnerabilities Potentially Hinders Maintainability

### 2020-05-29 paper/icsme20.tex
spellcheck
fixes

Novel tools have been built to 
utomatically detect software vulnerabilities (e.g., FindBugs, Infer and more). 
Despite these efforts, improving software security is not a trivial task and requires 
implementing patches that might affect software maintainability. 
Our suspicion is that some of these patches may have a negative 
impact on the software maintainability and, possibly, even be the cause of the 
introduction of new vulnerabilities---harming software reliability. Therefore, 
in this paper, we present an empirical study on the impact of patches of 
vulnerabilities on software maintenance across open-source software.
%
% Software maintainability can also be defined as 
% the property of software that reveals how easily the software system can be 
% maintained~\cite{IEEEComputerSociety:2014:GSE:2616205} or recover after a failure. 
% Hedegus et al.
% ($2018$)~\cite{HEGEDUS2018313}, analyzed the difference of maintainability between
% groups of refactored elements and non-refactored and conclude that the source
% code elements subjected to refactorings had significantly lower maintainability
% than the elements not modified.
% There is still little information about the impact of patching vulnerabilities
% in open-source software maintainability. \textit{Are developers affecting software
% maintainability when trying to improve software security?}, it is what we want to
% investigate.
%
As ISO does not provide any specific guidelines/formulas to calculate 
maintainability, we resort to Software Improvement Group (SIG)'s web-based source 
code analysis service Better Code Hub (BCH) to compute the software compliance 
with a set of $10$ guidelines/metrics to produce quality software 
based in ISO/IEC $25010$~\cite{Visser:2016:OREILLY}. 
There are other well-known standards and models that have been proposed
to increase software security: Common Criteria~\cite{common:2009} which received
negative criticism regarding the costs associated and poor technical evaluation;
the OWASP Application Security Verification Criteria~\cite{oswap:2009} which is
focused only on web applications, and a model proposed by Xu et al. ($2013$)
~\cite{6616351} for rating software security (arguably, it was one of the
first steps taken by SIG to introduce security on their maintainability model). 
Nevertheless, our study uses BCH to provide an assessment of maintainability in 
software for the following reasons: BCH integrates a broad number of code metrics; and, 
code metrics were empirically validated in previous work~\cite{Bijlsma:2012:FIR:2317098.2317124, 8530041, cruz2019energyoriented}.

@@

\noindent
where $G$ is the group of maintainability guidelines from BCH
(Table~\ref{tab:guidelines}) and $v$ is the version of the software under
evaluation. $M(v) < 0$ indicates that version $v$ is violating (some of) the
guidelines, while $M(v) > 0$ indicates that version $v$ is following
the BCH guidelines. The maintenance for the guideline $g$, $M_g$ for a given version
of a project is computed as the summation of the compliance with the 
maintainability guideline for the given severity level (medium, high, and very high).
The compliance for a severity level
is calculated based on previous work, which calculates the number of lines of code that comply and not comply
with the guideline at a given severity level~\cite{cruz2019energyoriented}. In our analysis, we compute 
the difference of maintainability between the security commit ($v_{s}$) and its parent 
commit ($v_{s-1}$), as illustrated in Figure~\ref{fig:commit}. Thus, we can determine 
which patches had a positive, negative or null impact on the project maintainability.

@@

This work presents an empirical study on the impact of $969$ security
patches on the maintainability of $260$ open-source projects. We leveraged
Better Code Hub reports to calculate maintainability based on a model proposed in 
previous work~\cite{Olivari:2018, cruz2019energyoriented}. Results show evidence of a 
trade-off between security and maintainability ($41.9\%$ of the cases), i.e., 
developers may be hindering software maintainability or even introduces 
new vulnerabilities while patching older ones. We also observe that some guidelines 
and programming languages are more likely to be affected than others. This study
aims to stress that changes to codebases while patching 
vulnerabilities need to be performed with extra care (e.g., integrating tools
for patch risk assessment in the CI/CD pipelines), computer science
curricula needs to be updated and more secure programming languages are necessary.

As future work, this study may be extended in several directions: 
investigate which guidelines affect most the maintainability per
weakness; check if vulnerability patches are followed by refactoring
commits and how much time does it take to do it; 
expand our methodology with other software quality properties; 
validate these findings with closed/private
software; and, expand this analysis to other quality standards.

>>>>

Novel tools have been built to automatically detect software vulnerabilities (e.g., FindBugs, Infer and more). 
Despite these efforts, improving software security is not a trivial task and requires 
implementing patches that might affect software maintainability. 
Our suspicion is that some of these patches may have a negative 
impact on the software maintainability and, possibly, even be the cause of the 
introduction of new vulnerabilities---harming software reliability and introducing technical debt. Therefore, 
in this paper, we present an empirical study on the impact of patches of 
vulnerabilities on software maintenance across open-source software.
%
% Software maintainability can also be defined as 
% the property of software that reveals how easily the software system can be 
% maintained~\cite{IEEEComputerSociety:2014:GSE:2616205} or recover after a failure. 
% Hedegus et al.
% ($2018$)~\cite{HEGEDUS2018313}, analyzed the difference of maintainability between
% groups of refactored elements and non-refactored and conclude that the source
% code elements subjected to refactorings had significantly lower maintainability
% than the elements not modified.
% There is still little information about the impact of patching vulnerabilities
% in open-source software maintainability. \textit{Are developers affecting software
% maintainability when trying to improve software security?}, it is what we want to
% investigate.
%
As ISO does not provide any specific guidelines/formulas to calculate 
maintainability, we resort to Software Improvement Group (SIG)'s web-based source 
code analysis service Better Code Hub (BCH) to compute the software compliance 
with a set of $10$ guidelines/metrics to produce quality software 
based in ISO/IEC $25010$~\cite{Visser:2016:OREILLY}. 
There are other well-known standards and models that have been proposed
to increase software security: Common Criteria~\cite{common:2009} which received
negative criticism regarding the costs associated and poor technical evaluation;
the OWASP Application Security Verification Criteria~\cite{oswap:2009} which is
focused only on web applications, and a model proposed by Xu et al. ($2013$)
~\cite{6616351} for rating software security (arguably, it was one of the
first steps taken by SIG to introduce security on their maintainability model). 
Nevertheless, our study uses BCH to provide an assessment of maintainability in 
software for the following reasons: BCH integrates a broad number of code metrics; and, 
code metrics were empirically validated in previous work~\cite{Bijlsma:2012:FIR:2317098.2317124, 8530041, cruz2019energyoriented}.

@@

\noindent
where $G$ is the group of maintainability guidelines from BCH
(Table~\ref{tab:guidelines}) and $v$ is the version of the software under
evaluation. $M(v) < 0$ indicates that version $v$ is violating (some of) the
guidelines, while $M(v) > 0$ indicates that version $v$ is following
the BCH guidelines. The maintenance for the guideline $g$, $M_g$ for a given version
of a project is computed as the summation of the compliance with the 
maintainability guideline for the given severity level (medium, high, and very high).
The compliance for a severity level
is calculated based on previous work, which calculates the number of lines of code that comply and not comply
with the guideline at a given severity level~\cite{cruz2019energyoriented}. In our analysis, we compute 
the difference of maintainability between the security commit ($v_{s}$) and its parent 
commit ($v_{s-1}$), as illustrated in Figure~\ref{fig:commit}. Thus, we can determine 
which patches had a positive, negative, or null impact on the project maintainability.

@@

This work presents an empirical study on the impact of $969$ security
patches on the maintainability of $260$ open-source projects. We leveraged
Better Code Hub reports to calculate maintainability based on a model proposed in 
previous work~\cite{Olivari:2018, cruz2019energyoriented}. Results show evidence of a 
trade-off between security and maintainability ($41.9\%$ of security patches yielded a negative impact), i.e., 
developers may be hindering software maintainability or even introduces 
new vulnerabilities while patching older ones. We also observe that some guidelines 
and programming languages are more likely to be affected than others. The implications 
of our study are that changes to codebases while patching 
vulnerabilities need to be performed with extra care (e.g., integrating tools
for patch risk assessment in the CI/CD pipelines); computer science
curricula needs to be updated; and, more secure programming languages are necessary.

As future work, the study extended in several directions: 
investigate which guidelines affect most the maintainability per
weakness; check if vulnerability patches are followed by refactoring
commits and how much time does it take to do it; 
expand our methodology with other software quality properties; 
validate these findings with closed/private
software; and, expand this analysis to other quality standards.
